---
title: "Week 4 R Notebook"
output: html_notebook
---

Getting and Cleaning Data Week 4 Quiz

Question 1 -- microdata and using strsplit()

first download and read data

```{r}
housingUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"

download.file(housingUrl, destfile = "housing.csv")

dateDownloaded <- date()

housingData <- read.table("housing.csv", sep = ",", header = TRUE)
```

now time to use strsplit()

```{r}
splitNames = strsplit(names(housingData), "wgtp")


answer1 = splitNames[[123]]

print(answer1)
```

Above, using strsplit() will remove the split paramater.

Compare with 123 element of housingData:
```{r}
housingData[123]

#Alternatively, view column 123 of data

View(housingData[123])
```

Question 2 -- remove comma and average GDP data

download and read data


```{r}
gdpData <- read.csv("gdp.csv") # this reads in 330 observations, reading in the original CSV file i downloaded

cleanGDP <- gsub(",", "", gdpData[,5]) #as above, this cleans out all rows from the 5th column, as character vector

mean(as.numeric(cleanGDP[1:194]), na.rm = TRUE) #this returns the answer
```

Question 3 -- using grep() to find countries that start with "United"

```{r}
q3 <- grep("^United",gdpData$X.2)

# this creates variable q3 that matches cases that BEGINS with "united" 
# a total of 3 observations (5, 10, 36) start with United

q3b <- grep("*United",gdpData$X.2)

# this would return ANY cases that have United

q3b2 <- grep("Rep.$",gdpData$X.2)

# this would return cases that END in "Rep."

```


Question 4 -- matching data; how many country's fiscal year ends in June?

```{r}
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv", destfile = "Q4GDPdata.csv")

Q4GDPdata <- read.csv("Q4GDPdata.csv") # this reads in more than 190 rows (330 observations, including many empty cells)

download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv", destfile = "Q4Eddata.csv") 

Q4Eddata <- read.csv("Q4Eddata.csv") # 234 observations, 31 variables (this contains the essential info for the question...no real need to match for answer)

Q4mergeddata <- merge(Q4Eddata, Q4GDPdata, by.x = "CountryCode", by.y = "X") 

  # will match different variable names, and keep the names from x

  #224 observations, 40 variables (only kept observations with country code, and all variables, minus the repeated     variable)

Q4answer <- length(grep("fiscal year end: june", tolower(Q4mergeddata$Special.Notes)))

print(Q4answer)

```


Question 5 -- quantmod package and values collected in 2012 / on Mondays in 2012

```{r}
install.packages("quantmod")  # download quantmod package

library(quantmod) # don't forget to load package!

amzn <- getSymbols("AMZN",auto.assign=FALSE)  

  # an xts (extentable time series) object from 2007-01-03 to 2018-06-11 containing 2880 observations, and 6 variables

sampleTimes <- index(amzn) # used index() to extract the 'index', that is, the dates (look like observations)

Q5a <- length(grep("^2012",sampleTimes))

print(Q5a)

library(lubridate) # you'll need lubridate to work with dates...

Q5b <- sum(wday(as.Date(sampleTimes[grep("^2012",sampleTimes)]))=="2")


  #summing up the extracted Mondays that match "^2012"" from sampleTimes

print(Q5b)

```

